\sc Cells at War: The Playfulness of Game-Based Learning

### 2023-04-12 thesis_template/paper0904.tex
DIscussion for paper finished

Aside from the benefits that game-based learning has for students utilizing the tool, the educators and students involved in the game design also reap massive benefits. There has been a rising interest in the realm of higher education about "students as partners" and co-created learning and teaching \citep{bibid}

\section{Future Directions}

>>>>

The Cells at War project was rooted in an interdisciplinary approach in which biology students, artists, programmers and game designers came together to share their expertise in their related fields. This approach benefited students in many ways, one of which was having students move from a more teacher centred approach to learning, to an instructional approach in which the students and educators co-create and co-learn together across a variety of disciplines. Benjamin Franklin’s old adage of “Tell me and I forget, teach me and I may remember, involve me and I learn” seems to have some merit as what the project served was to foster a learning environment in which students become empowered, engaged and self-motivated when working and learning alongside their teachers. Universities are typically structured in discipline-centred silos, which encourage discipline specific research. Many benefits arise however when individuals from a variety of disciplines come together \citep{snow201026,martin2020value}. Interdisciplinary teams are effective when they address complex, real-life problems and incorporate different perspectives \citep{beers2006common, choi2007multidisciplinarity, fischer2005knowledge, massey2006crossing, martin2020value}. Further, interdisciplinary projects encourage face-to-face interactions, develop a shared understanding of the problems they are attempting to address, and achieve effective team dynamics by building trusting, open environments that are not based on hierarchy \citep{stokols2005vivo, fischer2005knowledge, bruusgaard2010we, martin2020value}. Cells at War is a game based learning project in which students engage in experiential learning. Students working in interdisciplinary settings are reported to benefit from experiential learning as this leads to critical self-reflection, the development of transferable skills, and contributes to new ideas \citep{gemmel2009comprehensive, martin2020value}. Students at McMaster University and George Brown College were involved in a symbiotic process in which educators and students with different perspectives, abilities, and ideas were able to combine and apply their strengths and diverse knowledge and skills to build an educational video game that served as an engaging learning tool.

Aside from the benefits of engaging in an interdisciplinary approach, Cells at War is a highly engaging instructional tool developed by students, for students. The success of the development of the video game rests in educators and students working together in the game design process. This collaborative approach reaps mutual benefits and rewards. There has been a rising interest in the realm of higher education about "students as partners" and co-created learning and teaching \citep{felten2014engaging, dunne2016design, mercer2017systematic, bovill2020co}. As pedagogy in higher education has evolved, co-creation of learning has created more inclusive classrooms where students are able to build positive relationships with teachers, and students with students \citep{bovill2020co}. As the McMaster and George Brown team engaged in the development process of the video game, what emerged was a strong student voice in which students had a say in what they were learning and how to approach it. The dialogue that emerged and the exchange of ideas not only served to create an innovative instructional tool but, reinforced and consolidated the many skills and concepts that students from each discipline were required to learn. Greater student-teacher interaction leads to students’ academic success, good academic performance, higher educational aspirations, personal and intellectual development, student satisfaction, and enhanced motivation \citep{bovill2020co}.

The benefits to students engaging in the learning process are many. Of equal importance is the benefit for educators engaged in this process. Using novel approaches to improve student learning is an ongoing consideration for all educators. The process in which Cells at War was developed, provided teachers with an alternate way to engage and motivate students through best practices. In addition, teachers were able to provide feedback to students, which also supported them in informing their next point of instruction.

%\section{Future Directions}
